Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-08-31 Manuscript/manuscript.Rmd
First paragraph edits

The advent of chromosome conformation capture (3C) sequencing technologies, and its successor Hi-C, have revealed the spatial organization of the human genome  [@lieberman2009comprehensive]. The genome has been shown to form tight, compact folds in order to fit inside the nucleus. This folding does not occur randomly; in fact, the genome exhibits a hierarchical 3D structure [@dixon2012topological]. Among this hierarchy are distict compartments where the genome folds and interacts with itself more frequently termed topologically associating domains (TADs). TADs have garnered much interest in recent years given that they are heavily linked to gene expression [@dixon2015chromatin]. Moreover, changing the 3D structure of the DNA, causing disruptions in the boundaries that demarcate TADs, can lead to adverse outcomes and diseases like cancer [@hnisz2016activation; @Flavahan2016; @song2017ctcf]. It has been thought that the collection of particular functional genomic elements may contribute to the formation of TADs. Namely, the CCCTC-binding factor (CTCF) has been shown to be enriched at TAD boundaries [@ghirlando2016ctcf]. Additional transcription factors including SMC1, SMC3, and RAD21 have also been shown to colocalize near TAD boundaries forming a cohesin ring under the proposed loop extrusion model, whereby loops are formed during interphase [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self]. Likewise, chromatin has been shown to exhibit certain modifications at TAD boundaries, evidenced by the frequent observation of the H3K36me3 histone mark [@dixon2012topological]. The growing number of publically available genomic annotation data and their distinct patterns prompt the use of computational approaches toward TAD boundary prediction.

>>>>

The advent of chromosome conformation capture (3C) sequencing technologies, and its successor Hi-C, have revealed the spatial organization of the human genome [??? So what?] [@lieberman2009comprehensive]. The genome has been shown to form tight, compact folds in order to fit inside the nucleus. [??? So what?] This folding does not occur randomly; in fact, the genome exhibits a hierarchical 3D structure [@dixon2012topological]. [??? So what?] [??? All the above is pointless babbling, rewrite]

[??? This is very vague:  gene expression regulation [@dixon2015chromatin]

Among them, Topologically Associated Domains [??? THIS IS THE RIGHT SPELLING!] (TADs) have attracted most attention given their prominent role in constraining enhancer-promoter communication [@Lupianez:2015aa; @Franke:2016aa] and might be related to genome stability [@Canela:2017aa].  Disruption of TAD boundaries can lead to developmental disorders [@lupianez2016breaking], cancer [@hnisz2016activation; @taberlay2016three] and other disorders [@Sun:2018aa; @Meaburn:2007aa]. [??? This level of clarity is required]

It has been thought that the collection of particular functional genomic elements may contribute to the formation of TADs. Namely, the CCCTC-binding factor (CTCF) has been shown to be enriched at TAD boundaries [@ghirlando2016ctcf]. Additional transcription factors including SMC1, SMC3, and RAD21 have also been shown to colocalize near TAD boundaries forming a cohesin ring under the proposed loop extrusion model, whereby loops are formed during interphase [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self]. Likewise, chromatin has been shown to exhibit certain modifications at TAD boundaries, evidenced by the frequent observation of the H3K36me3 histone mark [@dixon2012topological]. The growing number of publically available genomic annotation data and their distinct patterns prompt the use of computational approaches toward TAD boundary prediction.
